---
output: github_document
---

<!-- badges: start -->
[![R-CMD-check](https://github.com/soylentOrange/sdarr/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/soylentOrange/sdarr/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# sdarr

The sdarr-package provides a R implementation of the SDAR algorithm for Slope Determination by Analysis of Residuals as standardized in [ASTM E3076-18](https://doi.org/10.1520/E3076-18).  
It allows for automated and objective linear-fitting of mechanical test-data. See a detailed description of the original algorithm in the [NIST Technical Note 2050 by E. Lucon](https://doi.org/10.6028/NIST.TN.2050).  
  
As the original algorithm heavily uses linear regressions, a faster version was implemented here, which finds the optimum region for linear fitting by random sub-sampling.

## Installation

You can install the development version of sdarr from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("soylentOrange/sdarr")
```

## Examples

### sdarr

Basic example of the sdarr-function on a synthetic data set.
The synthetic data set is based on the properties of aluminium (Al 6060 T66) as given in the 
[Metallic Material Properties Development and Standardization (MMPDS) Handbook](https://ntrl.ntis.gov/NTRL/dashboard/searchResults/titleDetail/PB2003106632.xhtml).
A toe-region and a non-zero intercept are added to make the test data less boring.  
  
The sdarr-function analyzes the data and will give a small report as a message.
It should confirm the Young's-modulus of 68 GPa and an intercept of 10 MPa.

```{r example-sdarr}
library(sdarr)

# Synthesize a test record resembling Al 6060 T66
# (Values from MMPDS Handbook, with a toe region added)
Al_6060_T66 <- synthesize_test_data(slope = 68000, yield.y = 160, 
                                    ultimate.y = 215, ultimate.x = 0.091,
                                    offset = 10,
                                    toe.initial.y = 3, toe.max.y = 15,
                                    toe.initial.slope = 34000)

# Analyze the test record
Al_6060_T66.result <- sdarr(Al_6060_T66, x = strain, y = stress, 
                            verbose = "r", showPlots = "r")

```

### sdarr.lazy

Basic example of the lazy variant of the sdarr-function on a synthetic data set.
The synthetic data set is based on the properties of aluminium (Al 6060 T66) as given in the 
[Metallic Material Properties Development and Standardization (MMPDS) Handbook](https://ntrl.ntis.gov/NTRL/dashboard/searchResults/titleDetail/PB2003106632.xhtml).
A toe-region and a non-zero intercept are added to make the test data less boring.  
  
The sdarr.lazy-function analyzes the data and will give a small report as a message.
It should confirm the Young's-modulus of 68 GPa and an intercept of 10 MPa.
As the data set is rather short, enforce random sub-sampling by setting enforce_subsampling to TRUE.

To make use of multi-core processing, configure [furrr](https://furrr.futureverse.org/) to use a multisession strategy.

```{r example-sdarr.lazy}
library(sdarr)

# setup multisession calculations with a maximum of 8 cores (or however many cores are available...)
library(furrr)
plan(multisession, workers = min(c(parallelly::availableCores(), 8)))

# Synthesize a test record resembling Al 6060 T66
# (Values from MMPDS Handbook, with a toe region added)
Al_6060_T66 <- synthesize_test_data(slope = 68000, yield.y = 160, 
                                    ultimate.y = 215, ultimate.x = 0.091,
                                    offset = 10,
                                    toe.initial.y = 3, toe.max.y = 15,
                                    toe.initial.slope = 34000)

# Analyze the test record
Al_6060_T66.result.lazy <- sdarr.lazy(Al_6060_T66, x = strain, y = stress,
                                      enforce_subsampling = T,
                                      verbose = "r", showPlots = "r")

```

## Acknowledgements

The sdarr-package was created for the analysis of mechanical test data for the project LOBio, which is funded by the German ministry of education and research (BMBF) under grant number 13XP5174C.
