---
output: github_document
---

<!-- badges: start -->
[![R-CMD-check](https://github.com/soylentOrange/sdarr/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/soylentOrange/sdarr/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# sdarr

The sdarr-package provides a R implementation of the SDAR algorithm for Slope Determination by Analysis of Residuals as standardized in [ASTM E3076-18](https://doi.org/10.1520/E3076-18).  
It allows for automated and objective linear-fitting of mechanical test-data. See a detailed description of the original algorithm in the [NIST Technical Note 2050 by E. Lucon](https://doi.org/10.6028/NIST.TN.2050).  
  
As the original algorithm heavily uses linear regressions, a faster version was implemented here, which finds the optimum region for linear fitting by random sub-sampling.

## Installation

You can install the development version of sdarr from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("soylentOrange/sdarr")
```

## Examples

### sdarr

Basic example of the sdarr-function on a synthetic data set, which is based on the properties of aluminium (Al 6060 T66) as given in the 
[Metallic Material Properties Development and Standardization (MMPDS) Handbook](https://ntrl.ntis.gov/NTRL/dashboard/searchResults/titleDetail/PB2003106632.xhtml).
A toe-region and a non-zero intercept are added to make the test data less boring.  
  
The sdarr-function analyzes the data and will give a small report as a message.
It should confirm the Young's-modulus of 68 GPa and an intercept of 10 MPa.
To make use of multi-core processing, configure [furrr](https://furrr.futureverse.org/) to use a multisession strategy.

```{r example-sdarr}
library(sdarr)

# setup multisession calculations with a maximum of 8 cores 
# (or however many cores are available...)
future::plan(future::multisession, 
             workers = min(c(parallelly::availableCores(), 8)))

# Synthesize a test record resembling Al 6060 T66
# (Values from MMPDS Handbook, with a toe region added)
Al_6060_T66 <- synthesize_test_data(slope = 68000, yield.y = 160, 
                                    ultimate.y = 215, ultimate.x = 0.091,
                                    offset = 10,
                                    toe.initial.y = 3, toe.max.y = 15,
                                    toe.initial.slope = 34000)

# Analyze the test record
Al_6060_T66.result <- sdarr(Al_6060_T66, x = strain, y = stress, 
                            verbose = "r", showPlots = "r")

```

### sdarr.lazy

Basic example of the lazy variant of the sdarr-function on a synthetic data set, which is based on the properties of aluminium (Al 6060 T66) as given in the 
[Metallic Material Properties Development and Standardization (MMPDS) Handbook](https://ntrl.ntis.gov/NTRL/dashboard/searchResults/titleDetail/PB2003106632.xhtml).
A toe-region and a non-zero intercept are added to make the test data less boring.  
  
The sdarr.lazy-function analyzes the data and will give a small report as a message.
It should confirm the Young's-modulus of 68 GPa and an intercept of 10 MPa.
As the data set is rather short, enforce random sub-sampling by setting enforce_subsampling to TRUE.

To make use of multi-core processing, configure [furrr](https://furrr.futureverse.org/) to use a multisession strategy.

```{r example-sdarr.lazy}
library(sdarr)

# setup multisession calculations with a maximum of 8 cores 
# (or however many cores are available...)
future::plan(future::multisession, 
             workers = min(c(parallelly::availableCores(), 8)))

# Synthesize a test record resembling Al 6060 T66
# (Values from MMPDS Handbook, with a toe region added)
Al_6060_T66 <- synthesize_test_data(slope = 68000, yield.y = 160, 
                                    ultimate.y = 215, ultimate.x = 0.091,
                                    offset = 10,
                                    toe.initial.y = 3, toe.max.y = 15,
                                    toe.initial.slope = 34000)

# Analyze the test record (after setting a random seed)
set.seed(50041180)
Al_6060_T66.result.lazy <- sdarr.lazy(Al_6060_T66, x = strain, y = stress,
                                      enforce_subsampling = T,
                                      verbose = "r", showPlots = "r")

```

### plot functions

The sdarr-function will create diagnostic plots throughout calculations, which will only be shown when requested (i.e. set showPlots = "all). To have a plot drawn later, you can call the provided plot-functions from the results, when you set savePlots = TRUE.  

The plots are crated functions (see [carrier](https://github.com/r-lib/carrier)), so you can easily tap the environment of the function to convert it to eg. a [ggplot2-graphic](https://ggplot2.tidyverse.org/).

```{r example-plot-fun}
library(sdarr)
# satisfy pipe addiction...
library(magrittr)
# make nice and shiny graphics withh ggplot2...
library(ggplot2)

# setup multisession calculations with a maximum of 8 cores 
# (or however many cores are available...)
future::plan(future::multisession, 
             workers = min(c(parallelly::availableCores(), 8)))

# Synthesize a test record resembling Al 6060 T66
# (Values from MMPDS Handbook, with a toe region added)
Al_6060_T66 <- synthesize_test_data(slope = 68000, yield.y = 160, 
                                    ultimate.y = 215, ultimate.x = 0.091,
                                    offset = 10,
                                    toe.initial.y = 3, toe.max.y = 15,
                                    toe.initial.slope = 34000)

# Analyze the test record using the standard algorithm
Al_6060_T66.result <- sdarr(Al_6060_T66, x = strain, y = stress,
                            verbose = "n", showPlots = "n", savePlots = TRUE)

# show plot of final fit using the plot function from the result
Al_6060_T66.result$plots$final.fit()

# plot fit using ggplot2 
{
  # tap the environment of the crated plot-function
  plot.env <- rlang::fn_env(Al_6060_T66.result$plots$final.fit)
  
  # get data and labels
  plot.data <- plot.env$plot.data
  plot.main <- plot.env$plot.main
  plot.xlab <- plot.env$plot.xlab
  plot.ylab <- plot.env$plot.ylab
  y.data.max <- plot.data$y.data %>% max()
  
  # create the ggplot2
  plot.data %>% ggplot(aes(x = x.data, y = y.data, color = "Al_6060_T66")) +
    geom_line() +
    geom_line(data = plot.data %>% 
                dplyr::filter(y.fit <= y.data.max),
              aes(x = x.data, y = y.fit, color = "fit (SDAR)")) +
    theme_bw() +
    labs(title = plot.main,
       x = plot.xlab,
       y = plot.ylab)
}

```

## Acknowledgements

The sdarr-package was created for the analysis of mechanical test data for the project LOBio, which is funded by the German ministry of education and research (BMBF) under grant number 13XP5174C.
