---
output: github_document
---

<!-- badges: start -->
[![R-CMD-check](https://github.com/soylentOrange/sdarr/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/soylentOrange/sdarr/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# sdarr

The sdarr-package provides a R implementation of the algorithm for Slope Determination by Analysis of Residuals (SDAR) as standardized in [ASTM E3076-18](https://doi.org/10.1520/E3076-18).  
It allows for automated and objective linear-fitting of mechanical test-data. See a detailed description of the algorithm in the [NIST Technical Note 2050 by E. Lucon](https://doi.org/10.6028/NIST.TN.2050) or in [Graham & Adler (2011)](https://doi.org/10.1520/JTE103038).  
  
As the SDAR-algorithm, implemented in `sdar()`, heavily uses linear regressions, a faster version `sdar.lazy()` was implemented, which finds the optimum region for the final linear regression by random sub-sampling within the normalized range of the test-data.
  
## Installation

You can install the latest release version of sdarr from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("soylentOrange/sdarr@*release")
```

You can install the development version of sdarr from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("soylentOrange/sdarr")
```

## Examples

### Standard SDAR-algorithm

A basic example of using `sdar()` on a synthetic data set, which is based on the properties of aluminium (Al 6060 T66) as given in the 
[Metallic Material Properties Development and Standardization (MMPDS) Handbook](https://ntrl.ntis.gov/NTRL/dashboard/searchResults/titleDetail/PB2003106632.xhtml).
A toe-region and a non-zero intercept are added to make the test data less boring.  
  
`sdar()` analyzes the data and will give a small report as a message.
It should confirm the Young's-modulus of 68 GPa and an intercept of 10 MPa.
To make use of multi-core processing, configure [furrr](https://furrr.futureverse.org/) to use a multisession strategy.

```{r example-sdar}
library(sdarr)

# setup multisession calculations with a maximum of 8 cores
# (or however many cores are available...)
future::plan(future::multisession,
  workers = min(c(parallelly::availableCores(), 8))
)

# Synthesize a test record resembling Al 6060 T66
# (Values from MMPDS Handbook, with a toe region added)
Al_6060_T66 <- synthesize_test_data(
  slope = 68000, yield.y = 160,
  ultimate.y = 215, ultimate.x = 0.091,
  offset = 10,
  toe.start.y = 3, toe.end.y = 15,
  toe.start.slope = 13600
)

# Analyze the test record
Al_6060_T66.result <- sdar(Al_6060_T66,
  x = strain, y = stress, plotFun = TRUE
)
```

### Random sub-sampling modification of the SDAR-algorithm

A basic example of `sdar.lazy()`, a random sub-sampling modification of the SDAR-algorithm on a synthetic data set, which is based on the properties of aluminium (Al 6060 T66) as given in the 
[Metallic Material Properties Development and Standardization (MMPDS) Handbook](https://ntrl.ntis.gov/NTRL/dashboard/searchResults/titleDetail/PB2003106632.xhtml).
A toe-region and a non-zero intercept are added to make the test data less boring (see above).  
  
`sdar.lazy()` analyzes the data for the optimum size of the fitting region via random sub-sampling. It will give a small report as a message after finding the optimum fit.
It should confirm the Young's-modulus of 68 GPa and an intercept of 10 MPa.
As the synthetic data set is noise-free (except for quantization-noise), only one random sub-sampling run will do.

To make use of multi-core processing, configure [furrr](https://furrr.futureverse.org/) to use a multisession strategy (see above).

```{r example-sdar.lazy}
# set a random seed
set.seed(50041180)

# Analyze the test record
# (with enforced random sub-sampling)
Al_6060_T66.result.lazy <- sdar.lazy(Al_6060_T66,
  x = strain, y = stress, plot = FALSE,
  plotFun = TRUE, n.fit = 1
)
```

### Plot Functions

`sdar()` and `sdar.lazy()` will create diagnostic plots throughout calculations, which will only be shown when requested (i.e. set `plotFun = TRUE` for obtaining a [crated](https://github.com/r-lib/carrier) plot-function of the final fit, or `plotFun.all = TRUE` for all additional diagnostic plots). To have a plot drawn later, you can call the provided plot-function from the results, when you set `plotFun = TRUE` (or `plotFun.all = TRUE`).  

The plot-functions are [crated](https://github.com/r-lib/carrier), so you can easily tap their environment to convert it into e.g. a [ggplot2-graphic](https://ggplot2.tidyverse.org/).

```{r example-plot-fun}
# show plot of final fit using the plot function from the result (see above)
Al_6060_T66.result.lazy$plots$final.fit()
```

```{r example-plot-fun-gg}
# satisfy pipe addiction...
library(magrittr)
# make nice and shiny graphics withh ggplot2...
library(ggplot2)

# plot the final fit using ggplot2
Al_6060_T66.result.lazy %>%
  {
    # tap the environment of the crated plot-function
    plot.env <- rlang::fn_env(.$plots$final.fit)

    # get data and labels
    plot.data <- plot.env$plot.data
    plot.main <- plot.env$plot.main
    plot.xlab <- plot.env$plot.xlab
    plot.ylab <- plot.env$plot.ylab
    plot.y.data.max <- plot.data$y.data %>% max()
    plot.y.lowerBound <- plot.env$y.lowerBound
    plot.y.upperBound <- plot.env$y.upperBound

    # create the ggplot2
    plot.data %>% ggplot(aes(x = x.data, y = y.data, color = "Test data\n(Al 6060 T66)")) +
      geom_line() +
      geom_line(
        data = plot.data %>%
          dplyr::filter(y.fit <= plot.y.data.max),
        aes(x = x.data, y = y.fit, color = "fit (sdar.lazy)")
      ) +
      geom_hline(
        aes(
          color = "fit range",
          yintercept = plot.y.lowerBound
        ),
        linetype = "dashed",
        show.legend = TRUE
      ) +
      geom_hline(
        aes(
          color = "fit range",
          yintercept = plot.y.upperBound
        ),
        linetype = "dashed",
        show.legend = TRUE
      ) +
      theme_bw() +
      labs(
        title = plot.main,
        x = plot.xlab,
        y = plot.ylab,
        caption = paste0(
          "Result of the random sub-sampling SDAR-algorithm:\n\nFinal Slope: ",
          round(.$sdar$finalSlope / 1000, 1), " GPa\nTrue Offset: ",
          round(.$sdar$trueIntercept, 1), " MPa\n\nFit Range: ",
          round(plot.y.lowerBound, 1), " MPa - ",
          round(plot.y.upperBound, 1), " MPa"
        )
      )
  }
```

# Acknowledgements

The sdarr-package was created for the analysis of mechanical test data for the project LOBio, which is funded by the German ministry of education and research (BMBF) under grant number 13XP5174C.
