---
title: "Validation against Reference Data Set"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
```

```{r setup, message = FALSE, warning = FALSE, include = FALSE}
# attach required packages
library(sdarr)
library(magrittr)
library(ggplot2)
library(plotly, include.only = c("ggplotly", "layout", "config"))

# setup multisession calculations with a maximum of 8 cores
# (or however many cores are available...)
future::plan(future::multisession,
  workers = min(c(parallelly::availableCores(), 8))
)
```

Though no exhaustive validation has been conducted, the Reference Data Set *E3076-DS1(2018) File 01* mentioned in [ASTM E3076-18](https://doi.org/10.1520/E3076-18) as *...available for download from ASTM...* has been used to check the results of this implementation. The Reference Data Set *E3076-DS1(2018) File 01* was also analyzed as benchmark in [NIST Technical Note 2050 by E. Lucon](https://doi.org/10.6028/NIST.TN.2050) with minor differences to the expected results listed in the standard.
  
  
## Reference Data Set 

A plot of the Reference Data Set (after downloading, unpacking and labeling) for validation of the Slope Determination by Analysis of Residuals (SDAR) algorithm is given below.

```{r Reference_Data_Set, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
# Get Data from Reference Data Set E3076-DS1(2018)
get_DS1.data <- function() {
  # satisfy pipe addiction...
  `%>%` <- magrittr::`%>%`

  # download file from the Internet into the temp-folder
  Reference_Data_Set.zip <- tempfile(fileext = ".zip")
  utils::download.file(
    "https://web.archive.org/web/20190905162007/https://www.astm.org/COMMIT/E3076-DS1(2018).zip",
    destfile = Reference_Data_Set.zip)

  # Extract the Zip Archive
  Reference_Data_Set.file <- utils::unzip(Reference_Data_Set.zip,
    exdir = tempdir())

  # read Data Set from unzipped file
  E3076_DS1_2018.rowwise <- readr::read_csv(Reference_Data_Set.file,
    col_names = FALSE,
    trim_ws = FALSE,
    show_col_types = FALSE) %>%
    dplyr::mutate(rownum = dplyr::row_number())

  # find start of the data section
  data.startIdx <- E3076_DS1_2018.rowwise %>%
    dplyr::mutate(data.start = stringr::str_detect(.data$X1, stringr::coll("<Data>"))) %>%
    dplyr::filter(data.start) %>% {
      .$rownum %>% as.numeric()
    }

  # find end of the data section
  data.endIdx <- E3076_DS1_2018.rowwise %>%
    dplyr::mutate(data.end = stringr::str_detect(.data$X1, stringr::coll("</Data>"))) %>%
    dplyr::filter(data.end) %>% {
      .$rownum %>% as.numeric()
    }

  # read the names and units from test record
  E3076_DS1_2018.names <- readr::read_table(Reference_Data_Set.file,
    show_col_types = FALSE,
    skip = data.startIdx + 3,
    n_max = 1)

  # read the data-points from the test record
  E3076_DS1_2018.data <- readr::read_table(Reference_Data_Set.file,
    col_names = FALSE,
    show_col_types = FALSE,
    skip = data.startIdx + 5,
    n_max = data.endIdx - data.startIdx - 5) %>%
    magrittr::set_names(E3076_DS1_2018.names %>% names()) %>%
    labelled::set_variable_labels(.labels = E3076_DS1_2018.names) %>%
    magrittr::set_attr("spec", NULL)

  # remove temporary files
  unlink(Reference_Data_Set.file)
  unlink(Reference_Data_Set.zip)

  # return the read and parsed data
  return(E3076_DS1_2018.data)
}

# Get Data from Reference Data Set E3076-DS1(2018)
E3076_DS1_2018 <- get_DS1.data()
```

```{r plot-Reference_Data_Set, echo=FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.width=6}
# plot of the data
E3076_DS1_2018 %>% {
  ggplot(data = ., aes(x = extensometer, y = force,
                       color = "Reference Data Set\n(E3076_DS1_2018)")) +
    geom_line() +
    theme_bw() +
    labs(title = "Plot of the Reference Data Set",
         x = "extensometer (in mm)",
         y = "force (in kN)")
} %>%
  ggplotly(dynamicTicks = TRUE, originalData = FALSE) %>%
  config(displaylogo = FALSE, showSendToCloud = FALSE)
```

# Validation of the standard SDAR-algorithm (sdar) {#link_to_standard_sdar}

```{r sdar_check, fig.height=4, fig.width=6}
# Run the algorithm
sdar_result <- E3076_DS1_2018 %>%
  sdar(extensometer, force,
    plot = FALSE, plotFun = TRUE)
```

```{r sdar_check_plot, echo=FALSE, out.width = "100%"}
# show the plot of the final fit
sdar_result %>% {
  # tap the environment of the crated plot-function
  plot.env <- rlang::fn_env(.$plots$final.fit)

  # get data and labels
  plot.data <- plot.env$plot.data
  plot.main <- plot.env$plot.main
  plot.xlab <- plot.env$plot.xlab
  plot.ylab <- plot.env$plot.ylab
  y.data.max <- plot.data$y.data %>% max()
  plot.y.lowerBound <- plot.env$y.lowerBound
  plot.y.upperBound <- plot.env$y.upperBound

  # create the ggplot2
  plot.data %>% ggplot(aes(x = x.data,
                           y = y.data,
                           color = "Reference Data Set\n(E3076_DS1_2018)")) +
    geom_line() +
    geom_line(data = plot.data %>%
                dplyr::filter(y.fit <= y.data.max),
              aes(x = x.data, y = y.fit, 
                  color = "fit (sdar)")) +
    geom_hline(aes(color = "fit range",
                   yintercept = plot.y.lowerBound),
               linetype = "dashed", show.legend = TRUE) +
    geom_hline(aes(color = "fit range",
                   yintercept = plot.y.upperBound),
               linetype = "dashed",show.legend = TRUE) +
    theme_bw() +
    labs(title = plot.main, x = plot.xlab, y = plot.ylab,
         caption = paste0("Result of the SDAR-algorithm:",
                          "\n\nFinal Slope: ",
                          round(.$sdar$finalSlope, 2), " kN/mm",
                          "\nTrue Intercept: ",
                          round(.$sdar$trueIntercept, 5), " MPa","
                          \n\nFit Range: ",
                          round(plot.y.lowerBound, 3), " kN - ",
                          round(plot.y.upperBound, 3), " kN"))
} %>% 
  plot()
```

## Validation Result of the standard SDAR-algorithm (sdar)

The results (see the [verbose messages](#link_to_standard_sdar) and table 1) of the analysis using `sdar()` are practically identical to the results given in [ASTM E3076-18](https://doi.org/10.1520/E3076-18) and perfectly match the results from [NIST Technical Note 2050 by E. Lucon](https://doi.org/10.6028/NIST.TN.2050) (except for omitting any rounding). As stated in [NIST Technical Note 2050 by E. Lucon](https://doi.org/10.6028/NIST.TN.2050) *...The few small discrepancies are most likely attributable to rounding errors...*.

```{r, message=FALSE, warning=FALSE, echo=FALSE, tab.cap="Table 1: Result comparison for the standard SDAR-algorithm"}

# use published results and new for sdar
sdar_result.table <- tibble::tibble("Parameter/result" = c("Relative x-resolution",
                                                           "Relative y-resolution",
                                                           "Relative x-noise",
                                                           "Relative y-noise",
                                                           "1st Quartile relative residual slope",
                                                           "4th Quartile relative residual slope",
                                                           "Relative Fit Range",
                                                           "Final Slope (in kN/mm)",
                                                           "True Intercept (in kN)",
                                                           "Lower y bound of fit range (in kN)",
                                                           "Upper y bound of fit range (in kN)"),
                                    "ASTM E3076" = c("0.67",
                                                     "0.33",
                                                     "0.325",
                                                     "0.144",
                                                     "0.366",
                                                     "0.205",
                                                     "0.92565",
                                                     "328.10",
                                                     "-0.03761",
                                                     "5.457",
                                                     "14.36"),
                                    "Excel spreadsheet" = c("0.67",
                                                           "0.33",
                                                           "0.325",
                                                           "0.144",
                                                           "0.366",
                                                           "0.205",
                                                           "0.92562",
                                                           "328.11",
                                                           "-0.03761",
                                                           "5.457",
                                                           "14.36"),
                                    "sdar" = c(paste0(round(sdar_result$Data_Quality_Metrics$digitalResolution$x$Relative_resolution, 2)),
                                               paste0(round(sdar_result$Data_Quality_Metrics$digitalResolution$y$Relative_resolution, 2)),
                                               paste0(round(sdar_result$Data_Quality_Metrics$Noise$x$Relative_noise, 3)),
                                               paste0(round(sdar_result$Data_Quality_Metrics$Noise$y$Relative_noise, 3)),
                                               paste0(round(sdar_result$Fit_Quality_Metrics$Curvature$first_quartile$relative_residual_slope, 3)),
                                               paste0(round(sdar_result$Fit_Quality_Metrics$Curvature$fourth_quartile$relative_residual_slope, 3)),
                                               paste0(round(sdar_result$Fit_Quality_Metrics$Fit_range$relative_fit_range, 5)),
                                               paste0(round(sdar_result$sdar$finalSlope %>% as.numeric(), 2)),
                                               paste0(round(sdar_result$sdar$trueIntercept %>% as.numeric(), 5)),
                                               paste0(round(sdar_result$sdar$y.lowerBound %>% as.numeric(), 3)),
                                               paste0(round(sdar_result$sdar$y.upperBound %>% as.numeric(), 3))))

sdar_result.table %>% 
  kableExtra::kbl(align = c("l", "c", "c", "c"), escape = FALSE,
                  caption = "Table 1: Result comparison for the standard SDAR-algorithm") %>%
  kableExtra::kable_styling(bootstrap_options = c("striped")) %>%
  kableExtra::row_spec(0, extra_css = "vertical-align:top;")
```

# Validation of the Random Sub-sampling Modification of the SDAR-algorithm (sdar_lazy) {#link_to_lazy_sdar}

```{r sdar.lazy_check}
# Run the algorithm (after setting a random seed)
set.seed(50041180)
# As the data set is rather short, the algorithm might refuse to use
# the random sub-sampling as it will require more computational effort.
# Enforce by setting enforce_subsampling = TRUE.

lazy_result <- E3076_DS1_2018 %>%
  sdar_lazy(extensometer, force,
    plot = FALSE, plotFun = TRUE,
    enforce_subsampling = TRUE)
```

```{r sdar.lazy_check_plot, echo=FALSE, out.width = "100%"}
# show the plot of the final fit
lazy_result %>% {
  # tap the environment of the crated plot-function
  plot.env <- rlang::fn_env(.$plots$final.fit)

  # get data and labels
  plot.data <- plot.env$plot.data
  plot.main <- plot.env$plot.main
  plot.xlab <- plot.env$plot.xlab
  plot.ylab <- plot.env$plot.ylab
  y.data.max <- plot.data$y.data %>% max()
  plot.y.lowerBound <- plot.env$y.lowerBound
  plot.y.upperBound <- plot.env$y.upperBound

  # create the ggplot2
  plot.data %>%
    ggplot(aes(x = x.data, y = y.data, 
               color = "Reference Data Set\n(E3076_DS1_2018)")) +
    geom_line() +
    geom_line(data = plot.data %>%
                dplyr::filter(y.fit <= y.data.max),
              aes(x = x.data, y = y.fit, 
                  color = "fit (sdar_lazy)")) +
    geom_hline(aes(color = "fit range",
                   yintercept = plot.y.lowerBound),
               linetype = "dashed", show.legend = TRUE) +
    geom_hline(aes(color = "fit range",
                   yintercept = plot.y.upperBound),
               linetype = "dashed",show.legend = TRUE) +
    theme_bw() +
    labs(title = plot.main, x = plot.xlab, y = plot.ylab,
         caption = paste0("Result of the random sub-sampling SDAR-algorithm:",
                          "\n\nFinal Slope: ",
                          round(.$sdar$finalSlope, 2), " kN/mm",
                          "\nTrue Intercept: ",
                          round(.$sdar$trueIntercept, 5), " MPa","
                          \n\nFit Range: ",
                          round(plot.y.lowerBound, 3), " kN - ",
                          round(plot.y.upperBound, 3), " kN"))
} %>% 
  plot()
```

## Validation Result of the Random Sub-sampling Modification of the SDAR-algorithm (sdar_lazy)

At first glimpse, the results (see the [verbose messages](#link_to_lazy_sdar) and table 2) of the analysis using the random sub-sampling variant of the SDAR-algorithm `sdar_lazy()` are not identical to the results given in [ASTM E3076-18](https://doi.org/10.1520/E3076-18) or the results from [NIST Technical Note 2050 by E. Lucon](https://doi.org/10.6028/NIST.TN.2050).  
  
However, when considering the 95% confidence interval of the final fitting (not mentioned in the ASTM E3076, but additionally calculated from the final linear regression of the data in `sdar()` and `sdar_lazy()`). The final slope of the lazy variant (`r paste0(round(lazy_result$sdar$finalSlope %>% as.numeric(), 2), " kN/mm")`) is well in line with the standard variant (95%-CI: `r paste0("[", round(sdar_result$sdar$finalSlope.conf.low %>% as.numeric(), 2), " kn/mm, ", round(sdar_result$sdar$finalSlope.conf.high %>% as.numeric(), 2), " kN/mm]")`). Similarly, the true intercept of the lazy variant (`r paste0(round(lazy_result$sdar$trueIntercept %>% as.numeric() * 1000, 1), " N")`) is well in line with the standard variant (95%-CI: `r paste0("[", round(sdar_result$sdar$trueIntercept.conf.low %>% as.numeric() * 1000, 1), " N, ", round(sdar_result$sdar$trueIntercept.conf.high %>% as.numeric() * 1000, 1), " N]")`).  
As the fitting range is estimated prior to finding a linear regression in `sdar_lazy()`, there is a minor difference in the y-range of the final fits that is resulting in slightly deviating final results. Comparing the regression curves of `sdar_lazy()` and `sdar()` in the plot below (plotted with the 95%-CI from `sdar()`), you see that there is practically no difference (zoom in ...a lot... to see a difference).

```{r, message=FALSE, warning=FALSE, echo=FALSE, tab.cap="Table 2: Result comparison for the random sub-sampling SDAR-algorithm"}

# insert new results for sdar_lazy
lazy_result.table <- sdar_result.table %>% 
  dplyr::select(- sdar) %>% 
  dplyr::bind_cols(tibble::tibble("sdar_lazy" = c(paste0(round(lazy_result$Data_Quality_Metrics$digitalResolution$x$Relative_resolution, 2)),
                                                  paste0(round(lazy_result$Data_Quality_Metrics$digitalResolution$y$Relative_resolution, 2)),
                                                  paste0(round(lazy_result$Data_Quality_Metrics$Noise$x$Relative_noise, 3)),
                                                  paste0(round(lazy_result$Data_Quality_Metrics$Noise$y$Relative_noise, 3)),
                                                  paste0(round(lazy_result$Fit_Quality_Metrics$Curvature$first_quartile$relative_residual_slope, 3)),
                                                  paste0(round(lazy_result$Fit_Quality_Metrics$Curvature$fourth_quartile$relative_residual_slope, 3)),
                                                  paste0(round(lazy_result$Fit_Quality_Metrics$Fit_range$relative_fit_range, 5)),
                                                  paste0(round(lazy_result$sdar$finalSlope %>% as.numeric(), 2)),
                                                  paste0(round(lazy_result$sdar$trueIntercept %>% as.numeric(), 5)),
                                                  paste0(round(lazy_result$sdar$y.lowerBound %>% as.numeric(), 3)),
                                                  paste0(round(lazy_result$sdar$y.upperBound %>% as.numeric(), 3)))))

lazy_result.table %>% 
  kableExtra::kbl(align = c("l", "c", "c", "c"), escape = FALSE,
                  caption = "Table 2: Result comparison for the random sub-sampling SDAR-algorithm") %>%
  kableExtra::kable_styling(bootstrap_options = c("striped")) %>%
  kableExtra::row_spec(0, extra_css = "vertical-align:top;")
```

```{r sdar.lazy_comparison, echo=FALSE, fig.height=4, fig.width=6}
# plot fit (with 95% CI) and lazy fit against original test record
{
  # tap the environment of the crated plot-functions
  plot.env <- rlang::fn_env(sdar_result$plots$final.fit)
  plot.env.lazy <- rlang::fn_env(lazy_result$plots$final.fit)

  # get data and labels
  plot.data <- plot.env$plot.data %>%
    dplyr::select(x.data, y.data)
  plot.main <- plot.env$plot.main
  plot.xlab <- plot.env$plot.xlab
  plot.ylab <- plot.env$plot.ylab

  # get fit ranges
  x.lowerBound <- sdar_result$sdar$x.lowerBound
  x.upperBound <- sdar_result$sdar$x.upperBound
  x.lowerBound.lazy <- lazy_result$sdar$x.lowerBound
  x.upperBound.lazy <- lazy_result$sdar$x.upperBound
  y.lowerBound <- sdar_result$sdar$y.lowerBound
  y.upperBound <- sdar_result$sdar$y.upperBound
  y.lowerBound.lazy <- lazy_result$sdar$y.lowerBound
  y.upperBound.lazy <- lazy_result$sdar$y.upperBound

  # re-fit for plotting
  y.data.prediction <- plot.data %>%
    dplyr::filter(dplyr::between(x.data, x.lowerBound, x.upperBound)) %>% {
      data <- .
      lm(y.data ~ x.data, data = data) %>%
        predict(newdata = dplyr::bind_rows(data.frame(x.data = 0),
                                           plot.data %>% dplyr::select(x.data)),
          interval = "confidence")
    }

  y.data.prediction.lazy <- plot.data %>%
    dplyr::filter(dplyr::between(x.data, 
                                 x.lowerBound.lazy, 
                                 x.upperBound.lazy)) %>% {
      data <- .
      lm(y.data ~ x.data, data = data) %>%
        predict(newdata = dplyr::bind_rows(data.frame(x.data = 0),
                                           plot.data %>%
                                             dplyr::select(x.data))) %>%
        as.data.frame() %>%
        set_names("fit.lazy")
    }

  # add an empty data point at 0 extension to see the difference in intercept
  plot.data <- data.frame(x.data = 0, y.data = NA) %>%
    dplyr::bind_rows(plot.data) %>%
    dplyr::bind_cols(y.data.prediction, y.data.prediction.lazy)

  y.data.max <- plot.data$y.data %>% max(na.rm = TRUE)

  ggplot(data = plot.data,
         aes(x = x.data, y = y.data, 
             color = "Reference Data Set\n(E3076_DS1_2018)")) +
    geom_line() +
    # geom_line(data = plot.data %>%
    #             dplyr::filter(dplyr::between(x.data, 
    #                                          x.lowerBound,x.upperBound)),
    #           aes(x = x.data, y = y.data,
    #               color = "fit-range (sdar)"),
    #           linewidth = 1.5, alpha = 0.5) +
    # geom_line(data = plot.data %>%
    #             dplyr::filter(dplyr::between(x.data, 
    #                                          x.lowerBound.lazy, 
    #                                          x.upperBound.lazy)),
    #           aes(x = x.data, y = y.data, 
    #               color = "fit-range (sdar_lazy)"),
    #           linewidth = 1.5, alpha = 0.5) +
    geom_line(data = plot.data %>%
                dplyr::filter(fit <= y.data.max),
              aes(x = x.data, y = fit, color = "fit (sdar)")) +
    geom_hline(aes(color = "fit range (sdar_lazy)",
                   yintercept = y.lowerBound.lazy),
               linetype = "dashed", show.legend = TRUE) +
    geom_hline(aes(color = "fit range (sdar_lazy)",
                   yintercept = y.upperBound.lazy),
               linetype = "dashed",show.legend = TRUE) +
    geom_hline(aes(color = "fit range (sdar)",
                   yintercept = y.lowerBound),
               linetype = "dashed", show.legend = TRUE) +
    geom_hline(aes(color = "fit range (sdar)",
                   yintercept = y.upperBound),
               linetype = "dashed",show.legend = TRUE) +
    geom_ribbon(data = plot.data %>%
                  dplyr::filter(fit <= y.data.max),
                aes(x = x.data, y = NULL, ymin = lwr, ymax = upr,
                    color = "fit (sdar)"),
                linewidth = 0, alpha = 0.25) +
    geom_line(data = plot.data %>%
                dplyr::filter(fit.lazy <= y.data.max),
              aes(x = x.data, y = fit.lazy, color = "fit (sdar_lazy)")) +
    theme_bw() +
    labs(title = "Final fit vs. final lazy fit",
         x = plot.xlab, y = plot.ylab)
} %>%
  ggplotly(dynamicTicks = TRUE, originalData = FALSE) %>%
  config(displaylogo = FALSE, showSendToCloud = FALSE)
```

# Conclusions

Even though this is more a benchmarking than a *real* validation, the results of `sdar()` are practically identical to the results listed in [ASTM E3076-18](https://doi.org/10.1520/E3076-18) and in [NIST Technical Note 2050 by E. Lucon](https://doi.org/10.6028/NIST.TN.2050). Considering the 95% confidence intervals of the fitting results, same seems true for the random sub-sampling modification used in `sdar_lazy()`.

## Final Remark
  
Even though the functions within this package have been tested, see the remark in the LICENSE...
  
THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
SOFTWARE.
