---
title: "Validation against Reference Data Set"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
```

`
```{r setup, message = FALSE, warning = FALSE, include = FALSE}
# attach required packages
library(sdarr)
library(magrittr)
library(stats)
library(ggplot2)
library(plotly)

# setup multisession calculations with a maximum of 8 cores
# (or however many cores are available...)
future::plan(future::multisession,
  workers = min(c(parallelly::availableCores(), 8))
)
```

Though no exhaustive validation has been conducted, the Reference Data Set *E3076-DS1(2018) File 01* mentioned in [ASTM E3076-18](https://doi.org/10.1520/E3076-18) as *...available for download from ASTM...* has been used to check the results of this implementation. The Reference Data Set *E3076-DS1(2018) File 01* was also analyzed as benchmark in [NIST Technical Note 2050 by E. Lucon](https://doi.org/10.6028/NIST.TN.2050) with minor differences to the expected results listed in the standard.
  
  
## Reference Data Set

A plot of the Reference Data Set (after downloading, unpacking and labeling) for validation of the Slope Determination by Analysis of Residuals (SDAR) algorithm is given below.

```{r Reference_Data_Set, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
# Get Data from Reference Data Set E3076-DS1(2018)
sdarr.get_DS1.data <- function() {
  # satisfy pipe addiction...
  `%>%` <- magrittr::`%>%`

  # download file from the Internet into the temp-folder
  Reference_Data_Set.zip <- tempfile(fileext = ".zip")
  utils::download.file(
    "https://web.archive.org/web/20190905162007/https://www.astm.org/COMMIT/E3076-DS1(2018).zip",
    destfile = Reference_Data_Set.zip
  )

  # Extract the Zip Archive
  Reference_Data_Set.file <- utils::unzip(Reference_Data_Set.zip,
    exdir = tempdir()
  )

  # read Data Set from unzipped file
  E3076_DS1_2018.rowwise <- readr::read_csv(Reference_Data_Set.file,
    col_names = FALSE,
    trim_ws = FALSE,
    show_col_types = FALSE
  ) %>%
    dplyr::mutate(rownum = dplyr::row_number())

  # find start of the data section
  data.startIdx <- E3076_DS1_2018.rowwise %>%
    dplyr::mutate(data.start = stringr::str_detect(.data$X1, stringr::coll("<Data>"))) %>%
    dplyr::filter(data.start) %>%
    {
      .$rownum %>% as.numeric()
    }

  # find end of the data section
  data.endIdx <- E3076_DS1_2018.rowwise %>%
    dplyr::mutate(data.end = stringr::str_detect(.data$X1, stringr::coll("</Data>"))) %>%
    dplyr::filter(data.end) %>%
    {
      .$rownum %>% as.numeric()
    }

  # read the names and units from test record
  E3076_DS1_2018.names <- readr::read_table(Reference_Data_Set.file,
    show_col_types = FALSE,
    skip = data.startIdx + 3,
    n_max = 1
  )

  # read the data-points from the test record
  E3076_DS1_2018.data <- readr::read_table(Reference_Data_Set.file,
    col_names = FALSE,
    show_col_types = FALSE,
    skip = data.startIdx + 5,
    n_max = data.endIdx - data.startIdx - 5
  ) %>%
    magrittr::set_names(E3076_DS1_2018.names %>% names()) %>%
    labelled::set_variable_labels(.labels = E3076_DS1_2018.names) %>%
    magrittr::set_attr("spec", NULL)

  # remove temporary files
  unlink(Reference_Data_Set.file)
  unlink(Reference_Data_Set.zip)

  # return the read and parsed data
  return(E3076_DS1_2018.data)
}

# Get Data from Reference Data Set E3076-DS1(2018)
E3076_DS1_2018 <- sdarr.get_DS1.data()
```

```{r plot-Reference_Data_Set, message=FALSE, warning=FALSE, fig.height=4, fig.width=6}
# plot of the data
E3076_DS1_2018 %>%
  {
    ggplot(data = ., aes(
      x = extensometer, y = force,
      color = "Reference Data Set\n(E3076_DS1_2018)"
    )) +
      geom_line() +
      theme_bw() +
      labs(
        title = "Plot of the Reference Data Set",
        x = "extensometer (in mm)",
        y = "force (in kN)"
      )
  } %>%
  ggplotly(dynamicTicks = TRUE, originalData = FALSE) %>%
  config(displaylogo = FALSE, showSendToCloud = FALSE)
```

# Validation of the standard SDAR-algorithm (sdar)

```{r sdar_check, fig.height=4, fig.width=6}
# Run the algorithm
sdar.results <- E3076_DS1_2018 %>%
  sdar(extensometer, force,
    plot = FALSE, plotFun = TRUE
  )

# show the plot of the final fit
sdar.results$plots$final.fit %>%
  {
    # tap the environment of the crated plot-function
    plot.env <- rlang::fn_env(.)

    # get data and labels
    plot.data <- plot.env$plot.data
    plot.main <- plot.env$plot.main
    plot.xlab <- plot.env$plot.xlab
    plot.ylab <- plot.env$plot.ylab
    y.data.max <- plot.data$y.data %>% max()

    # create the ggplot2
    plot.data %>% ggplot(aes(
      x = x.data,
      y = y.data,
      color = "Reference Data Set\n(E3076_DS1_2018)"
    )) +
      geom_line() +
      geom_line(
        data = plot.data %>%
          dplyr::filter(y.fit <= y.data.max),
        aes(x = x.data, y = y.fit, color = "fit (sdar)")
      ) +
      theme_bw() +
      labs(
        title = plot.main,
        x = plot.xlab,
        y = plot.ylab
      )
  } %>%
  ggplotly(dynamicTicks = TRUE, originalData = FALSE) %>%
  config(displaylogo = FALSE, showSendToCloud = FALSE)
```

## Validation Result of the standard SDAR-algorithm (sdar)

The results of the analysis using `sdar()` are practically identical to the results given in [ASTM E3076-18](https://doi.org/10.1520/E3076-18) and perfectly match the results from [NIST Technical Note 2050 by E. Lucon](https://doi.org/10.6028/NIST.TN.2050) (except for omitting any rounding). As stated in [NIST Technical Note 2050 by E. Lucon](https://doi.org/10.6028/NIST.TN.2050) *...The few small discrepancies are most likely attributable to rounding errors...*.

# Validation of the Random Sub-sampling Modification of the SDAR-algorithm (sdar.lazy)

```{r sdar.lazy_check, fig.height=4, fig.width=6}
# Run the algorithm (after setting a random seed)
# As the data set is rather short, the algorithm might refuse to use
# the random sub-sampling as it will require more computational effort.
# Enforce by setting enforce_subsampling = TRUE.
set.seed(50041180)
sdar.lazy.results <- E3076_DS1_2018 %>%
  sdar.lazy(extensometer, force,
    # cutoff_probability = 0.5,
    # n.fit = 5,
    # n.candidates = 5,
    # enforce_subsampling = TRUE,
    plot = FALSE, plotFun = TRUE
  )

# show the plot of the final fit
sdar.lazy.results$plots$final.fit %>%
  {
    # tap the environment of the crated plot-function
    plot.env <- rlang::fn_env(.)

    # get data and labels
    plot.data <- plot.env$plot.data
    plot.main <- plot.env$plot.main
    plot.xlab <- plot.env$plot.xlab
    plot.ylab <- plot.env$plot.ylab
    y.data.max <- plot.data$y.data %>% max()

    # create the ggplot2
    plot.data %>%
      ggplot(aes(x = x.data, y = y.data, color = "Reference Data Set\n(E3076_DS1_2018)")) +
      geom_line() +
      geom_line(
        data = plot.data %>%
          dplyr::filter(y.fit <= y.data.max),
        aes(
          x = x.data, y = y.fit,
          color = "fit (sdar.lazy)"
        )
      ) +
      theme_bw() +
      labs(
        title = plot.main,
        x = plot.xlab,
        y = plot.ylab
      )
  } %>%
  ggplotly(dynamicTicks = TRUE, originalData = FALSE) %>%
  config(displaylogo = FALSE, showSendToCloud = FALSE)
```

## Validation Result of the Random Sub-sampling Modification of the SDAR-algorithm (sdar.lazy)

At first glimpse, the results of the analysis using the random sub-sampling variant of the SDAR-algorithm `sdar.lazy()` are not identical to the results given in [ASTM E3076-18](https://doi.org/10.1520/E3076-18) or the results from [NIST Technical Note 2050 by E. Lucon](https://doi.org/10.6028/NIST.TN.2050) (except for omitting any rounding).  
However, when considering the 95% confidence intervals (additionally given with the result from `sdar()` and `sdar.lazy()`) obtained from the final fitting in the standard SDAR-algorithm, the results of the lazy variant are well within the 95%-CI of the standard variant. There is only a minor difference in the y-range of the final fits resulting in slightly different final results.  
Comparing the regression curves of `sdar.lazy()` and `sdar()` over the mechanical data, you see that there is practically no difference (zoom in ...a lot... to see a difference).

```{r sdar.lazy_comparison, fig.height=4, fig.width=6}
# plot fit (with 95% CI) and lazy fit against original test record
{
  # tap the environment of the crated plot-functions
  plot.env <- rlang::fn_env(sdar.results$plots$final.fit)
  plot.env.lazy <- rlang::fn_env(sdar.lazy.results$plots$final.fit)

  # get data and labels
  plot.data <- plot.env$plot.data %>%
    dplyr::select(x.data, y.data)
  plot.main <- plot.env$plot.main
  plot.xlab <- plot.env$plot.xlab
  plot.ylab <- plot.env$plot.ylab

  # get fit ranges
  x.lowerBound <- sdar.results$sdar$x.lowerBound
  x.upperBound <- sdar.results$sdar$x.upperBound
  x.lowerBound.lazy <- sdar.lazy.results$sdar$x.lowerBound
  x.upperBound.lazy <- sdar.lazy.results$sdar$x.upperBound

  # re-fit for plotting
  y.data.prediction <- plot.data %>%
    dplyr::filter(dplyr::between(x.data, x.lowerBound, x.upperBound)) %>%
    {
      data <- .
      lm(y.data ~ x.data, data = data) %>%
        predict(
          newdata = dplyr::bind_rows(
            data.frame(x.data = 0),
            plot.data %>% dplyr::select(x.data)
          ),
          interval = "confidence"
        )
    }

  y.data.prediction.lazy <- plot.data %>%
    dplyr::filter(dplyr::between(
      x.data,
      x.lowerBound.lazy,
      x.upperBound.lazy
    )) %>%
    {
      data <- .
      lm(y.data ~ x.data, data = data) %>%
        predict(newdata = dplyr::bind_rows(
          data.frame(x.data = 0),
          plot.data %>%
            dplyr::select(x.data)
        )) %>%
        as.data.frame() %>%
        set_names("fit.lazy")
    }

  # add an empty data point at 0 extension to see the difference in intercept
  plot.data <- data.frame(x.data = 0, y.data = NA) %>%
    dplyr::bind_rows(plot.data) %>%
    dplyr::bind_cols(
      y.data.prediction,
      y.data.prediction.lazy
    )

  y.data.max <- plot.data$y.data %>% max(na.rm = TRUE)

  ggplot(
    data = plot.data,
    aes(x = x.data, y = y.data, color = "Reference Data Set\n(E3076_DS1_2018)")
  ) +
    geom_line() +
    geom_line(
      data = plot.data %>%
        dplyr::filter(dplyr::between(
          x.data,
          x.lowerBound,
          x.upperBound
        )),
      aes(
        x = x.data, y = y.data,
        color = "fit-range (sdar)"
      ),
      linewidth = 1.5, alpha = 0.5
    ) +
    geom_line(
      data = plot.data %>%
        dplyr::filter(dplyr::between(
          x.data,
          x.lowerBound.lazy,
          x.upperBound.lazy
        )),
      aes(
        x = x.data, y = y.data,
        color = "fit-range (sdar.lazy)"
      ),
      linewidth = 1.5, alpha = 0.5
    ) +
    geom_line(
      data = plot.data %>%
        dplyr::filter(fit <= y.data.max),
      aes(x = x.data, y = fit, color = "fit (sdar)")
    ) +
    geom_ribbon(
      data = plot.data %>%
        dplyr::filter(fit <= y.data.max),
      aes(
        x = x.data, y = NULL, ymin = lwr, ymax = upr,
        color = "fit (sdar)"
      ),
      linewidth = 0, alpha = 0.25
    ) +
    geom_line(
      data = plot.data %>%
        dplyr::filter(fit.lazy <= y.data.max),
      aes(x = x.data, y = fit.lazy, color = "fit (sdar.lazy)")
    ) +
    theme_bw() +
    labs(
      title = "Final fit (with 95% CI) vs. final lazy fit",
      x = plot.xlab,
      y = plot.ylab
    )
} %>%
  ggplotly(dynamicTicks = TRUE, originalData = FALSE) %>%
  config(displaylogo = FALSE, showSendToCloud = FALSE)
```

# Conclusions

Even though this is more a benchmarking than a *real* validation, the results of `sdar()` are practically identical to the results listed in [ASTM E3076-18](https://doi.org/10.1520/E3076-18) and in [NIST Technical Note 2050 by E. Lucon](https://doi.org/10.6028/NIST.TN.2050). Considering the 95% confidence intervals of the fitting results, same seems true for the random sub-sampling modification used in `sdar.lazy()`.

## Final Remark
  
Even though the functions within this package have been tested, see the remark in the LICENSE...
  
THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
SOFTWARE.
